---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# pppindexr


Price and Quantity Indices for International Comparisons: The EKS, CCD, and the 
Geary-Khamis real income indices.

Rudimentary functionality and some unit tests are implemented.
## Installation

You can install the development version of pppindexr from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("ErikOSorensen/pppindexr")
```

## Example

Let us look at a 2x2 example (two countries of equal size, two goods) and define
some arbitrary prices and quantities (one row per country):

```{r example}
library(pppindexr)
ex1 <- data.frame(population=c(1,1),
                  p1=c(1,2), p2=c(1,1), 
                  q1=c(1,1), q2=c(2,2))
# EKS index and Gheary Khamis:
ex1$eks = eks(ex1[,2:3],ex1[,4:5])
ex1$gk  = gk( ex1[,2:3], ex1[,4:5], pop=ex1$pop)
```


