---
title: "Real income indices on data from 'Rationalizing the Penn World Table'"
author: "Erik Ø. Sørensen"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Rationalizing-the-Penn-World-Table}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
library(pppindexr)
library(tidyverse)
```

Peter Neary wrote an important paper in the AER (2004) comparing
different approaches to calculating real income indices using the International
Comparison Project (ICP) data -- the same data used in the Penn World Table (PWT). 
Neary proposes an improvement on the PWT use of the Geary-Khamis method for 
calculating real income. I use Neary's data to illustrate how to use the 
functions I provide.

# The data and implied indices.
Neary uses a set of 11 goods for 60 countries. I have compiled his data
in the `Neary2004` data set available with the `pppindexr` package. In addition
to prices and "quantites" (expenditure at US prices), there are country names and
population, all for the ICP baseline year 1980. 

I can use the implemented quantity index functions to add the indices
to the data. I scale all observations in the minimum (Egypt), the same
way that Geary does for his Table 1 and Figure 1.

First I extract the price and quantity matrices:
```{r}
P <- Neary2004 %>% dplyr::select(p1:p11) %>% as.matrix()
Q <- Neary2004 %>% dplyr::select(q1:q11) %>% as.matrix()
```

Then I add the indices to the Neary2004 dataset. I enter two different
Geary-Khamis numbers: one with and one without weighting by population
size.
```{r}
Neary2004$eks = eks(P,Q)
Neary2004$ccd = ccd(P,Q)
Neary2004$gk =  ygk(P,Q)
Neary2004$gk_pop =  ygk(P,Q, pop=Neary2004$pop1980)
Neary2004$gk = Neary2004$gk/Neary2004$gk[60]
Neary2004$gk_pop = Neary2004$gk_pop/Neary2004$gk_pop[60]
```

Now I list the data to compare it to Table 1 of Neary:

```{r}
Neary2004 %>% select(country, eks, ccd, gk, gk_pop) %>% knitr::kable(digits=3)
```

Inspecting the table, the `eks`, the `ccd`, and the `gk` columns replicate
exactly the corresponding columns in Table 1 of Neary (2004), 
The population weighted Geary Khamis (`gk_pop`) numbers are slightly different
from the ones reported by Neary (`gk`). This means
that Neary calculated Geary-Khamis quantities without population weighting,
which I believe is contrary to standard practice - it means that subdividing gdp by 
region within a country (with uniform within-country prices) will lead
to different world prices, and hence also different real income indices. 

I also show how to calculate the world prices:
```{r}
prices <- tibble(good_number = 1:11,
                 p_gk = pgk(P,Q),
                 p_gk_pop = pgk(P,Q, pop=Neary2004$pop1980))
prices %>%  knitr::kable(digits=3)
```


# References
- Peter J. Neary (2004). "Rationalizing the
  Penn World Table: True Multilateral Indices for International Comparisons
  of Real Income." *American Economic Review*, 94 (5): 1411-1428.
  [https://www.aeaweb.org/articles?id=10.1257/0002828043052286]


